Investigating Predictors for American Attitudes on Justifiability of Abortion

### 2022-06-20 Final/FinalReport.qmd
model eval done

The dataset used in this analysis observes attitudes on the
justifiability of abortion among respondents in in the United States
across six "waves" of the World Values Survey (1982-2011), which is
administered every few years and collects information about people's
values and beliefs worldwide, alongside basic demographic
characteristics. The data are collected via face-to-face interviews at
the respondents' homes. In this data set, the response variable,
justifiability of abortion, is a numerical measure on a scale of 1 to 10
on the individual person's attitude toward whether abortion is
justifiable or not. Individuals responded 1 for "abortion is never
justified" and 10 for "abortion is always justified.

Our research question is as follows: "Do an individual's political
ideology and their personal attitudes/preferences towards other issues,
such as the importance of religion in their life and their respect for
authority, among others, predict their attitude on the justifiability of
abortion?" We will attempt to answer this question using an EDA-informed
predictive model. Given that this issue has become highly polarized by
political party, we predict that liberal ideology and liberal-leaning
attitudes on other issues will correlate with belief that abortion is
more justified.

@@

\[3 line plots, displaying AIC, BIC and R\^2 value respectively, against
Fold number. Each plot should have the line graph for each of the 3
models we used)

```{r cv}
#| echo: false
set.seed(206)
folds <- vfold_cv(abortion_training, v = 10)
abortion_fit_rs <- abortion_wflow %>%
  fit_resamples(folds)

abortion_fit_rs2 <- abortion_wflow2 %>%
  fit_resamples(folds)

abortion_fit_rs3 <- abortion_wflow3 %>%
  fit_resamples(folds)
```

```{r cv-table}
#| echo: false
cv_metrics <- collect_metrics(abortion_fit_rs, summarize = FALSE) 
cv_metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) 
cv_metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE)

cv_metrics %>%
  mutate(.estimate = round(.estimate, 3)) %>%
  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
  kable(col.names = c("Fold", "RMSE", "R-squared"))

cv_metrics2 %>%
  mutate(.estimate = round(.estimate, 3)) %>%
  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
  kable(col.names = c("Fold", "RMSE", "R-squared"))

cv_metrics3 %>%
  mutate(.estimate = round(.estimate, 3)) %>%
  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
  kable(col.names = c("Fold", "RMSE", "R-squared"))
```

```{r aic-bic}
#| echo: false
glance(abortion_fit) %>% 
  select(AIC, BIC)

glance(abortion_fit2) %>% 
  select(AIC, BIC)

glance(abortion_fit3) %>% 
  select(AIC, BIC)
```

From the above AIC and BIC values, we have evidence that our chosen
model is the strongest model and best fit for the data compared to two
alternatives with fewer predictors. Our model has the lowest AIC and BIC
values, which are usually penalized for having more predictors, but in
this case the model is strong enough to overcome that penalty.

>>>>

The dataset used in this analysis observes attitudes on the
justifiability of abortion among respondents in in the United States
across six "waves" of the World Values Survey (1982-2011), which is
administered every few years and collects information about people's
values and beliefs worldwide, alongside basic demographic
characteristics. The data are a nationally representative sample of the
American population and are collected via face-to-face interviews at the
respondents' homes. In this data set, the response variable,
justifiability of abortion, is a numerical measure on a scale of 1 to 10
on the individual person's attitude toward whether abortion is
justifiable or not. Individuals responded 1 for "abortion is never
justified" and 10 for "abortion is always justified.

Our research question is as follows: "Do an individual's political
ideology and their personal attitudes/preferences towards other issues,
such as the importance of religion in their life and their respect for
authority, among others, predict their attitude on the justifiability of
abortion?" We will attempt to answer this question using an EDA-informed
predictive model. Because the data are nationally representative, we can
infer that all the observations are independent and the variables of
interest may show a linear relationship. Given that this issue has
become highly polarized by political party, we predict that liberal
ideology and liberal-leaning attitudes on other issues will correlate
with belief that abortion is more justified.

@@

```{r cv}
#| message: false
#| echo: false
set.seed(206)
folds <- vfold_cv(abortion_training, v = 10)

abortion_fit_rs <- abortion_wflow %>%
  fit_resamples(folds)
abortion_fit_rs2 <- abortion_wflow2 %>%
  fit_resamples(folds)
abortion_fit_rs3 <- abortion_wflow3 %>%
  fit_resamples(folds)

metrics1 <- collect_metrics(abortion_fit_rs, summarize = FALSE) %>%
  mutate(model = "Model 1")
metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) %>%
  mutate(model = "Model 2")
metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE) %>%
  mutate(model = "Model 3")

metrics <- bind_rows(metrics1, metrics2, metrics3) %>%
  arrange(.metric)

ggplot(metrics %>% filter(.metric == "rsq"),
	aes(x = id, y = .estimate, group = model, color = model)) +
    geom_point() +
    geom_line() +
    labs(title = "R-squared")
```

From the above visualization of R-squared values, we have evidence that
our chosen model is the strongest model and best fit for the data
compared to two alternatives with fewer predictors. Below we will
observe the AIC, BIC, and Adjusted R-squared values for each model as
well.

```{r aic-bic-adjrsq}
#| echo: false
abortion_table = matrix(c(glance(abortion_fit)$AIC,
                          glance(abortion_fit)$BIC,
                          glance(abortion_fit)$adj.r.squared,
                          glance(abortion_fit2)$AIC,
                          glance(abortion_fit2)$BIC, 
                          glance(abortion_fit2)$adj.r.squared, 
                          glance(abortion_fit3)$AIC,
                          glance(abortion_fit3)$BIC,
                          glance(abortion_fit3)$adj.r.squared), 
                        ncol = 3, byrow = TRUE)

colnames(abortion_table) = c('AIC','BIC','Adj. R-squared')
rownames(abortion_table) = c('Model 1','Model 2','Model 3')

final_table = as.table(abortion_table)
final_table %>%
  kable(digits = 3)
```

The table above further supports that Model 1 is our best fitting model.
